One exact and one random sampling method to assess the reliability of inferences from network enrichment analysis

### 2018-07-31 doc/genesetdp.tex
Equation -> Eq

\begin{equation}
p(s)=\frac{N(s)/2 +\sum_{s'=s+1}^{S} N(s')}{\sum_{s'=0}^{S} N(s')},
\label{eq:pval}
\end{equation}
where $S$ is the maximal score, $s$, that $Q$ genes can obtain from Equation~(\ref{eq:sum_ij}).

We can reformulate Equation~(\ref{eq:sum_ij}) by defining the number of links, $l_i=\sum_{j=1}^P x_{ij}$, which gives us a score,
\begin{equation}
s=\sum_{i=1}^Q l_i.
\label{eq:sum_i}
\end{equation}
Such number of links, $l_i$, can be pre-computed for all genes in $\mathcal{G}$. For later purposes we also define a mapping, $k_a=\sum_{\{i:l_i=a\}}1$, giving the number of genes having $a$ links to the investigated pathway.

@@

We first implemented a random sampling algorithm to assess $N(s)$. We randomly selected sets of $Q$ genes from $\mathcal{G}$. These sets scores were calculated with Equation~(\ref{eq:sum_i}). We counted the number of times, $F_B(s)$, a score of $s$ was obtained when sampling $B$ gene sets. We see that $F_B(s)$ will approximately have the same shape as $N(s)$ when selecting a large $B$, and we can hence calculate $p$~values by replacing $N$ with $F_B$ in Equation~(\ref{eq:pval}). We refer to this procedure as GeneSetMC.

@@

We implemented a Python program that reads network and pathway definition files and scores a query sets against a pathway according to Equation~(\ref{eq:sum_i}), using GeneSetDP and GeneSetMC described in the Algorithm section, that enabled us to assign $p$~values according to Equation~(\ref{eq:pval}). We downloaded pathway and network definitions from the BinoX's website and used the same network threshold as BinoX default ($0.7$).

>>>>

\begin{equation}
p(s)=\frac{N(s)/2 +\sum_{s'=s+1}^{S} N(s')}{\sum_{s'=0}^{S} N(s')},
\label{eq:pval}
\end{equation}
where $S$ is the maximal score, $s$, that $Q$ genes can obtain from Eq~(\ref{eq:sum_ij}).

We can reformulate Eq~(\ref{eq:sum_ij}) by defining the number of links, $l_i=\sum_{j=1}^P x_{ij}$, which gives us a score,
\begin{equation}
s=\sum_{i=1}^Q l_i.
\label{eq:sum_i}
\end{equation}
Such number of links, $l_i$, can be pre-computed for all genes in $\mathcal{G}$. For later purposes we also define a mapping, $k_a=\sum_{\{i:l_i=a\}}1$, giving the number of genes having $a$ links to the investigated pathway.

@@

We first implemented a random sampling algorithm to assess $N(s)$. We randomly selected sets of $Q$ genes from $\mathcal{G}$. These sets scores were calculated with Eq~(\ref{eq:sum_i}). We counted the number of times, $F_B(s)$, a score of $s$ was obtained when sampling $B$ gene sets. We see that $F_B(s)$ will approximately have the same shape as $N(s)$ when selecting a large $B$, and we can hence calculate $p$~values by replacing $N$ with $F_B$ in Eq~(\ref{eq:pval}). We refer to this procedure as GeneSetMC.

@@

We implemented a Python program that reads network and pathway definition files and scores a query sets against a pathway according to Eq~(\ref{eq:sum_i}), using GeneSetDP and GeneSetMC described in the Algorithm section, that enabled us to assign $p$~values according to Eq~(\ref{eq:pval}). We downloaded pathway and network definitions from the BinoX's website and used the same network threshold as BinoX default ($0.7$).
